---
title: "Strings and Factors"
output: github_document
---
this is the repository for data wrangling 2 part of P8105
```{r}
library(rvest)
library(p8105.datasets)
library(tidyverse)
library(httr)
```

## Lets do strings
```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
```

```{r}
str_replace(string_vec, "jeff", "Jeff")
```

For exact matches, you can designate matches at the beginning or end of a line.
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")
```

You can designate a list of characters that will count as a match.
```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Pp]umpkin")
```

You don’t have to list these; instead, you can provide a range of letters or numbers that count as a match.
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")

```

The character . matches anything.
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

Some characters are “special”. These include [ and ], ( and ), and .. If you want to search for these, you have to indicate they’re special using \. Unfortunately, \ is also special, so things get weird.
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```

## Factors
```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex

as.numeric(vec_sex)
```

```{r}
vec_sex = fct_relevel(vec_sex, "male")
vec_sex

as.numeric(vec_sex)

```

